Index measures for oak decline severity using phenotypic descriptors

### 2019-12-15 manuscript/manuscript.Rmd
updated lime DAI method. Response surface results complete. Added points to lime results section

Local Interpretable Model-Agnostic Explanations (LIME) were computed to identify descriptors contributing locally to case specific examples of each of the indexes [@ribeiro2016]. 
The lime R package v`r packageVersion('lime') %>% as.character()` was used with the default parameters [@pedersen2019].
A healthy, a moderate and a severe decline cases was used for the PDI. 
For each case, a tree with a score closest to 0, 0.5 and 1 respectively was selected.
For the DAI, a decline neutral case was selected with a score closes to 0.
Then for each of the AOD and COD syndromes, moderate and severe cases were selected with scores closes to 0.5 and 1 for AOD and -0.5 and -1 for COD.

@@

- Lime results

@@

- Top ranked descriptor index response surfaces
- Trends of key phenotypic descriptors within the PDI and DAI measures
- Relatively smaller trees associated with more severe decline.

These complex interactions can be investigated by simulating phenotypic data and generating the index values from random forest models that have been trained for each of the indexes, supervised using phenotypic descriptors from the 171 trees. 
The index values generated by the models from these simulated phenotypic data can then be visualised as response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 1 and 2 show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

Crown condition descriptors such as crown volume and crown condition contributed highly to the PDI. 
The phenotypic measures central to this were percentage missing crown and percentage crown transparency from which crown condition and crown volume are calculated. 
It can be seen in Figure 1 that as the percentage missing crown and percentage crown transparency increase and therefore crown condition and estimated crown volume decrease, the PDI increases. 
A much greater range of PDI values can be obtained from trees with relatively greater stature (e.g. greater crown radius, total height, lower crown height and diameter at breast height; Figures 1a & b), with relatively small stature trees having PDIs of 0.5 or greater even with 100% crown condition.

The DAI is a much more complex index with more descriptors contributing to it than the PDI such as bleed prevalence, diameter at breast height and crown condition descriptors. 
Simulated response surfaces for the DAI are shown in Figure 2. 
These response surfaces show circumstances under which bleed prevalence related descriptors (Figures 2a, b & c) and crown condition descriptors (Figures 2d, e & f) change for non-symptomatic AOD or COD trees (Figures 2a & d respectively), moderate AOD or COD (Figures 2b & e respectively) and severe AOD or COD (Figures 2c & f respectively). 
Increasing active bleed, staining size and frequency contributed to greater positive values of the DAI and more severe AOD. 
Also higher DAI values could be obtained from trees with greater stature (Figure 2c). 
Similarly to the PDI, trees with smaller stature and poorer crown condition give lower DAI values (Figures 2d, e & f) and therefore more severe COD.
The relationship of relative tree stature and the acuteness of the decline likely reflects the period of time under which these trees have been afflict. 
AOD occurs over a much shorter period compared COD and therefore has less time in which to inhibit growth of the tree as a whole.

>>>>

Local Interpretable Model-Agnostic Explanations (LIME) were computed to identify descriptors contributing locally to case specific examples of each of the indexes [@ribeiro2016]. 
The lime R package v`r packageVersion('lime') %>% as.character()` was used with the default parameters [@pedersen2019].
A healthy, a moderate and a severe decline cases was used for the PDI. 
For each case, a tree with a score closest to 0, 0.5 and 1 respectively was selected.
For the DAI cases were selected from trees with a PDI score greater than 0.5, 
A decline neutral case was selected with a score closest to 0.
Then for each of the AOD and COD syndromes, moderate and severe cases were selected with scores closes to 0.5 and 1 for AOD and -0.5 and -1 for COD.

@@

- Important descriptors identified model-wide aren't necessarily relevant to localised cases.

- Top descriptors for each case same as top descriptors in overall model importance lists for PDI.
- Decline is associated with a reduction in the size and condition of the crown

- Different descriptors locally compared to overall importance for DAI.
- COD trees have greater crown contact than AOD trees.
- Bleed prevalence not included in any local importances
- Descriptors relevant to context

@@

The complex descriptor interactions within the index models can be investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data can then be visualised as response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures # and # show the simulated index value response surfaces under a variety of different circumstances for key phenotypic descriptors that contribute highly to the PDI and DAI.

Crown condition descriptors such as crown volume and crown condition contributed highly to the PDI (Figure #). 
The phenotypic measures central to this were percentage missing crown and percentage crown transparency from which crown condition and crown volume are calculated. 
It can be seen in Figure # that as the percentage missing crown and percentage crown transparency increase and therefore crown condition and estimated crown volume decrease, the PDI increases. 
A much greater range of PDI values can be obtained from trees with relatively greater stature (e.g. greater crown radius, total height, lower crown height and diameter at breast height; Figures #a & b), with relatively small stature trees having PDIs of 0.5 or greater even with 100% crown condition.

The DAI is a much more complex index with more descriptors contributing to it than the PDI such as bleed prevalence, diameter at breast height and crown condition descriptors. 
Simulated response surfaces for the DAI are shown in Figure c. 
These response surfaces show circumstances under which bleed prevalence related descriptors (Figures #a, b & c) and crown condition descriptors (Figures dd, e & f) change for non-symptomatic AOD or COD trees (Figures #a & d respectively), moderate AOD or COD (Figures #b & e respectively) and severe AOD or COD (Figures #c & f respectively). 
Increasing active bleed, staining size and frequency contributed to greater positive values of the DAI and more severe AOD. 
Also higher DAI values could be obtained from trees with greater stature (Figure #c). 
Crown transparency and missing crown have no influence on the DAI for trees of neutral decline type (Figure #d).
Similarly to the PDI, trees with smaller stature and poorer crown condition give lower DAI values (Figures #e & f) and therefore more severe COD.
The relationship of relative tree stature and the acuteness of the decline likely reflects the period of time under which these trees have been afflict. 
AOD occurs over a much shorter period compared COD and therefore has less time in which to inhibit growth of the tree as a whole.
